Bridging the Gap Between Visual and Analytical Machine Learning Testing

### 2023-09-21 report/esecfse24.tex
feat(report): 2023-09-21 20:36 updates

\usepackage{listings}
\usepackage{graphicx}

@@

\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

@@

\section{Preliminaries}\label{sec:prelim}
\subsection{Related Work}\label{sec:related}
\subsection{Background}\label{sec:background}
\subsection{Internal Structure of Jupyter Notebooks}\ref{sec:nbformat}
% TODO we need to explain the json structure of jupyter notebooks, we reference this in the methodology to explain the github query.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\section{Methodology}\label{sec:method}
% TODO is there a scientific term in SE research to explain reading code in detail?
% NOTE: we looked for symmantic similarity between the visualisation and assertion code
This section presents the methodology used to conduct the empirical
study presented in Section~\ref{sec:intro}. Figure~\ref{fig:method}
presents an overview of all relevant steps used in this study. The
empirical analysis comprises of three distinct phases namely:
\textit{the data collection phase}, \textit{identification of related
visualisation and assertion pairs} and finally a \textit{complete
analysis of the notebooks} to identify the final set of 29
visualisation and assertion pairs.

@@

\subsection{Deeper Analysis of Notebooks}\ref{sec:deep-dive}
% describe the steps used for the deeper analysis of NBs
% - we read the NBs in top-down fashion
% - reading the introductory md cell at the top often gives the context for the contents of the NB (eg. what type of ML model? which part of the ML lifecycle is this relevant for?)
% - reading the md cell above visualisation to understand context/motivation of visualisation
% - reading the source code that produces the visualisation
% - reading md cell after visualisation to understand the author's interpretation/observations from visualisation
% - reading the assertion
% - we fould 29 unique patterns from the 90 NBs; we highlight the 5 most interesting patterns in results

@@

\subsection{Decision boundary of SVM and accuracy}

% % TODO try using the minipage environment
% \begin{table}
%   \centering
%   \caption{TODO}
%   \begin{tabular}[t]{p{0.5\textwidth} p{0.5\textwidth}}
%   \toprule
%   \textbf{Visualisation} &
%   \textbf{Assertion}\\
%   \midrule
%   \includegraphics[width=\linewidth]{../catalogue/select-04.png} &
%   \begin{verbatim}
%   assert accuracy_score(y_test, pred) > 0.95
%   \end{verbatim}\\
%   \includegraphics[width=\linewidth]{../catalogue/select-09.png} &
%   \begin{verbatim}
%   assert roc_auc_score(
%     y_test,
%     wv_model.predict_proba(X_test_wv)[:, 1]
%   ) > 0.92,
%     "something's wrong with your featurs"
%   \end{verbatim}\\
%   \bottomrule
%   \end{tabular}
%   \label{TODO}
% \end{table}

\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python]
    assert accuracy_score(y_test, pred) > 0.95
  \end{lstlisting}
  \begin{minipage}{0.5\textwidth}
    \includegraphics{../catalogue/select-04.png}
  \end{minipage}
\end{minipage}

>>>>

\usepackage{listings}

\lstdefinestyle{mystyle}{
  basicstyle=\ttfamily\footnotesize,
  breakatwhitespace=true,
  breaklines=true,
  keepspaces=true,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=2,
  captionpos=b
}
\lstset{style=mystyle}
\usepackage{graphicx}
\usepackage{caption}

@@

\begin{abstract}
% what is the problem?
% why is this an important problem to solve?
% what did we do?
% what did we find?

Testing ML systems is a highly interactive process which demands a
human-in-the-loop approach. In addition to writing tests for the code
base, practitioners are required to analyse and interpret several
visualisations using their domain expertise to validate if an ML system
satisfies the required set of functional and non-functional properties.
Visualisations are frequently used to qualitatively assess various
parts of an ML pipeline. However, implicit knowledge gained from
visualisations must be translated to explicit analytical tests that fail
when there is change in any component of the ML pipeline. We conduct an
empirical analysis of Jupyter notebooks to catalogue the state-of-the-
art mappings between ML visualisations and assertions. We mine Github to
collect Jupyter Notebooks that contain assertions written in Python. We
develop a novel methodology to identify 1764 notebooks which contain an
assertion below a visualisation. We manually analyse the 1.7K notebooks
and identify 551 notebooks where at least one pair of visualisation and
assertion are semantically related to one another. We further read and
understand 90 notebooks in depth to identify 26 qualitative and their
corresponding quantitative testing patterns. We showcase five such
patterns in this paper and plan to make the entire catalogue publicly
available in the near future.

\end{abstract}

@@

\section{Preliminaries}\label{sec:prelim}
\subsection{Related Work}\label{sec:related}
% summarise existing work in SE on computational NBs: bavishi2021vizsmith, pimentel2019large, quaranta2021kgtorrent
\subsection{Background}\label{sec:background}
\subsection{Internal Structure of Jupyter Notebooks}\ref{sec:nbformat}
% TODO we need to explain the json structure of jupyter notebooks, we reference this in the methodology to explain the github query.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\section{Methodology}\label{sec:method}
% TODO is there a scientific term in SE research to explain reading code in detail?
% NOTE we looked for semantic similarity between the visualisation and assertion code

This section presents the methodology used to conduct the empirical
study presented in Section~\ref{sec:intro}. Figure~\ref{fig:method}
presents an overview of all relevant steps used in this study. The
empirical analysis comprises of three distinct phases namely:
\textit{the data collection phase}, \textit{identification of related
visualisation and assertion pairs} and finally a \textit{complete
analysis of the notebooks} to identify the final set of 29
visualisation and assertion pairs.

@@

\subsection{Deeper Analysis of Notebooks}\label{sec:deep-dive}
% describe the steps used for the deeper analysis of NBs
% - we read the NBs in top-down fashion
% - reading the introductory md cell at the top often gives the context for the contents of the NB (eg. what type of ML model? which part of the ML lifecycle is this relevant for?)
% - reading the md cell above visualisation to understand context/motivation of visualisation
% - reading the source code that produces the visualisation
% - reading md cell after visualisation to understand the author's interpretation/observations from visualisation
% - reading the assertion
% - we fould 29 unique patterns from the 90 NBs; we highlight the 5 most interesting patterns in results

@@

\subsection{Decision boundary}\label{sec:svm}

% TODO try using captionof for the listing as well, so that the caption for the listing and figure align at the bottom
\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python, caption=Assertion on the accuracy of the ML model.]
assert accuracy_score(y_test, pred) > 0.95
  \end{lstlisting}
  \label{lst:svm}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-04.png}
  \captionof{figure}{Visualisation of the decision boundary of a SVM.}
  \label{fig:svm}
\end{minipage}

Figure~\ref{fig:svm} presents a visualisation of the decision boundaries of a Support Vector Machine (SVM). The author is visualising how the SVM is distinguishing between the two classes present in the dataset. Listing~\ref{lst:svm} shows the corresponding assertion defined below the visualisation.

The assertion is  
\subsection{ROC Curve}

\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python]
assert roc_auc_score(y_test, wv_model.predict_proba(X_test_wv)[:, 1] ) > 0.92, "something's wrong with your features"
  \end{lstlisting}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-09.png}
\end{minipage}

\subsection{Training loss}

\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python]
assert np.mean(history[:10]) >np.mean(history[-10:]), "RNN didn't converge"
  \end{lstlisting}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-12.png}
\end{minipage}

\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python]
# assert we actually descended at each step
for i in range(1, n_steps):
    assert loss_array_true_grad[i] - loss_array_true_grad[i-1]) <= 0
  \end{lstlisting}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-267.png}
\end{minipage}

\subsection{Coefficient of Determination}

\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python]
r2_gru = r2_score(y_test, y_pred)
assert r2_gru > 0.6
  \end{lstlisting}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-332a.png}
\end{minipage}

\subsection{Time-series Prediction}

\begin{minipage}{0.5\textwidth}
  \begin{lstlisting}[language=Python]
# Result of the wfc function for the best choice of s
assert np.isclose(best_wfv, 7685.546728971963, 0.001)
  \end{lstlisting}
\end{minipage}
\begin{minipage}{0.5\textwidth}
  \includegraphics[width=\linewidth]{../catalogue/select-117b.png}
\end{minipage}
